BACI of biomass change for VOCC source cells

### 2019-08-20 analysis/VOCC/4-vocc-BACI.Rmd
add gamm to BACI analysis

#### Make climate vectors for biannual changes between surveys
```{r echo=TRUE}
trend <- FALSE
start_time <- 2012
end_time <- 2015
temp_threshold <- c(0.1)
do_threshold <- c(0.1)
```

@@

Run all subsequent code...
```{r, include=FALSE}
library(dplyr)
library(ggplot2)
# library(gfplot)
# library(gfdata)
library(sdmTMB)
library(gfranges)
```

@@

if (climate == "temperature") {
  variable_names <- c("temp")
  min_thresholds <- c(Inf)
  delta_threshold <- temp_threshold
  optimal_values <- optimal_temp

@@

if (climate == "do") {
  variable_names <- c("do_est")
  max_thresholds <- c(Inf)
  delta_threshold <- -1 * (do_threshold)
  optimal_values <- optimal_do

@@

if (length(vectors[, 2]) == 0) {
  other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
  coefs <- other_coefs[1, ]
  coefs$var <- "NA"
  coefs$beta <- "NA"
  coefs$se <- "NA"
  coefs$model <- "NA"
  coefs$species <- spp
  coefs$lifestage <- "NA"
  coefs$climate <- climate
  coefs$min_thresholds <- min_string
  coefs$max_thresholds <- max_string
  coefs$optimum <- optimal_string
  coefs$start_time <- start_time
  coefs$timespan <- delta_t_total
  coefs$region <- region
  coefs$ssid_string <- ssid_string

  coefs <- rbind(other_coefs, coefs) %>% distinct()
  saveRDS(coefs, file = paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))

  stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
}
# glimpse(vocc)
# glimpse(vectors)
```

@@

```{r}
data <- vocc_by_sp[["adult"]]
#View(data)
data$log_density <-  log(data$density)

ggplot(data, aes(y=log_density, x=temp, colour=as.factor(after))) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C")

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)

ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)
```

```{r}
m <- lmerTest::lmer(log_density~after + cell_type + after:cell_type + std_temp + std_temp2 + (1|matchobs), data=data)
summary(m)
```

```{r}
sjPlot::plot_model(m, title = paste(species, "adult"))
```

@@

```{r}
idata <- vocc_by_sp[["imm"]]
#View(data)
idata$log_density <-  log(idata$density)

ggplot(idata, aes(y=log_density, x=temp, colour=as.factor(after))) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C")

ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)

ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)

ggplot(idata, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~after)
```

```{r}
mi <- lmerTest::lmer(log_density~after + cell_type + after:cell_type + std_temp + std_temp2 + (1|matchobs), data=idata)
summary(mi)
```

```{r}
sjPlot::plot_model(mi, title = paste(species, "immature"))
```

>>>>

#### Make climate vectors for biannual changes between surveys
```{r echo=TRUE}
trend <- FALSE
start_time <- 2012
end_time <- 2015
temp_threshold <- c(0.5)
do_threshold <- c(0.1)
```

@@

Run all subsequent code...
```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
# library(gfdata)
library(sdmTMB)
#library(itsadug)
library(gfranges)
```

@@

if (climate == "temperature") {
  variable_names <- c("temp")
  min_thresholds <- c(Inf)
  #delta_threshold <- temp_threshold
  optimal_values <- optimal_temp

@@

if (climate == "do") {
  variable_names <- c("do_est")
  max_thresholds <- c(Inf)
  #delta_threshold <- -1 * (do_threshold)
  optimal_values <- optimal_do

@@

# if (length(vectors[, 2]) == 0) {
#   other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
#   coefs <- other_coefs[1, ]
#   coefs$var <- "NA"
#   coefs$beta <- "NA"
#   coefs$se <- "NA"
#   coefs$model <- "NA"
#   coefs$species <- spp
#   coefs$lifestage <- "NA"
#   coefs$climate <- climate
#   coefs$min_thresholds <- min_string
#   coefs$max_thresholds <- max_string
#   coefs$optimum <- optimal_string
#   coefs$start_time <- start_time
#   coefs$timespan <- delta_t_total
#   coefs$region <- region
#   coefs$ssid_string <- ssid_string
# 
#   coefs <- rbind(other_coefs, coefs) %>% distinct()
#   saveRDS(coefs, file = paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
# 
#   stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
# }
# glimpse(vocc)
# glimpse(vectors)
```

@@

Prep data
```{r}
data <- vocc_by_sp[["adult"]]
#View(data)
data$log_density <-  log(data$density)
data$distance2 <- data$distance^2

data <- mutate(data, time = ifelse(after==1, "end", "begin"))
data$matchobs <- as.factor(data$matchobs)
data$origobs  <- as.factor(data$origobs)   
```

Plot data
```{r}
ggplot(data, aes(y=log_density, x=temp, colour=time)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C")

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)

ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)
```

Basic BACI
```{r}

m <- lmerTest::lmer(log_density~after + cell_type + after:cell_type + std_temp + std_temp2 + distance + distance2 + (1|icell) + (1|matchobs), data=data)
summary(m)
```

```{r}
sjPlot::plot_model(m, type = "std2", order.terms=c(1,2,7,5,6,3,4), title = paste(species, "adult"))
```

GAMM
```{r}
gam4 <- gamm4::gamm4(log_density~ time * cell_type + s(std_temp) + s(distance) + s(X,Y), 
  random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")

plot(gam4$gam,pages=1)

summary(gam4$gam) ## summary of gam
summary(gam4$mer) ## underlying mixed model
anova(gam4$gam)

# itsadug::plot_parametric(gam4$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")), 
#   parametricOnly = TRUE, main="") + scale_y_reverse()

plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
   parametricOnly = TRUE, main="", xlab ="log_density") 
```

@@

Prep data
```{r}
idata <- vocc_by_sp[["imm"]]
#View(data)
idata$log_density <-  log(idata$density)
idata$distance2 <- idata$distance^2

idata <- mutate(idata, time = ifelse(after==1, "end", "begin"))
idata$matchobs <- as.factor(idata$matchobs)
idata$origobs  <- as.factor(idata$origobs)   
```

Plot data
```{r}
ggplot(idata, aes(y=log_density, x=temp, colour=time)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C")

ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)

ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)

ggplot(idata, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)
```

Basic BACI
```{r}
mi <- lmerTest::lmer(log_density ~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + distance + distance2 + (1|icell) + (1|matchobs), data=idata)
summary(mi)
```

```{r}
sjPlot::plot_model(mi, type = "std2", title = paste(species, "immature"))
```

GAMM
```{r}
igam4 <- gamm4::gamm4(log_density~ time * cell_type + s(std_temp) + s(distance) + s(X,Y), 
  random = ~(1|matchobs) + (1|origobs), data=idata) # + s(X,Y, bs="re")

plot(igam4$gam,pages=1)

summary(igam4$gam) ## summary of gam
summary(igam4$mer) ## underlying mixed model
anova(igam4$gam)

# itsadug::plot_parametric(gam4$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")), 
#   parametricOnly = TRUE, main="") + scale_y_reverse()

plot_parametric(igam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
   parametricOnly = TRUE, main="", xlab ="log_density") 
```

Misc exploring gamm 
```{r eval=FALSE}
# mi <- mgcv::gamm(log_density~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + distance + distance2 + s(as.factor(idata$icell), bs="re") + s(as.factor(idata$matchobs), bs="re"), data=idata)
# summary(mi)
# 
# mi <- mgcv::gamm(log_density~ as.factor(after) * cell_type + s(std_temp) + s(distance) + s(as.factor(idata$matchobs), bs="re") , data=idata) # + s(X,Y, bs="re")
# summary(mi)
# mi
# 
# 
# idata$after <- as.factor(idata$after)
# idata$matchobs <- as.factor(idata$matchobs)
# 
# m4b <- gamm4::gamm4(log_density~ after * cell_type + s(std_temp) + s(distance) + #s(X) + s(Y) + 
#     s(X,Y), random = ~(1|matchobs), data=idata) # + s(X,Y, bs="re")
# 
# 
# plot(m4b$gam,pages=1)
# 
# summary(m4b$gam) ## summary of gam
# summary(m4b$mer) ## underlying mixed model
# anova(m4b$gam) 
# 
# gamtabs(m1)
# 
# #library(itsadug)
# plot_parametric(m4b$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")))
# 
# library(mgcv)
```
